---
title: "Analysis of the PlantGrowth Dataset"
author: "Paul Schmidt"
date: today
format: docx
---

## Introduction

This report analzes the `PlantGrowth` dataset, which contains results from an experiment comparing plant yields under three different conditions: a control group (`ctrl`) and two treatment groups (`trt1` and `trt2`). The goal is to determin whether there are statistically significant differences in dried plant weight between the groups.

## Data Overview

The dataset contains 30 observations across three groups with 10 plants each.

```{r}
data(PlantGrowth)

group_summary <- aggregate(
  weight ~ group,
  data = PlantGrowth,
  FUN = function(x) c(mean = mean(x), sd = sd(x), n = length(x))
)

summary_df <- data.frame(
  Group = group_summary$group,
  Mean = round(group_summary$weight[, "mean"], 2),
  SD = round(group_summary$weight[, "sd"], 2),
  n = group_summary$weight[, "n"]
)

knitr::kable(summary_df, caption = "Summary statistics by treatment group.")
```

## Visualization

```{r}
#| label: fig-boxplot
#| fig-cap: "Distribution of dried plant weight by treatment group."

library(ggplot2)

ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot(width = 0.5, outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.6) +
  labs(
    x = "Treatment Group",
    y = "Dried Weight [g]"
  ) +
  theme_minimal()
```

## Statistical Analysis

We use a one-way ANOVA to test whether the group means differ significantly, followed by Tukey's HSD test for pairwise comparisons.

```{r}
model <- aov(weight ~ group, data = PlantGrowth)
summary(model)
```

```{r}
TukeyHSD(model)
```

## Conclusion

The ANOVA indicates a marginally significant effect of treatment group on plant weight. Tukey's HSD test reveals that the difference between `trt2` and `trt1` is statistically significant, while neither treatment group differs significantly from the control.
